
```{r}
rm(list=ls())
```

#libraries
```{r}
library(dplyr)
```

## the Datasets
```{r}
waterData = read.csv("water_dataX.csv")
str(waterData)
```

## Cleaning Water data
```{r}
waterData = waterData %>%
  rename(DO=D.O...mg.l.,pH=PH,Conductivity=CONDUCTIVITY..µmhos.cm., BOD=B.O.D...mg.l., Nitrates=NITRATENAN.N..NITRITENANN..mg.l., FecalColiform=FECAL.COLIFORM..MPN.100ml., TotalColiform=TOTAL.COLIFORM..MPN.100ml.Mean)
waterData$Conductivity = as.numeric(waterData$Conductivity)
waterData$BOD = as.numeric(waterData$BOD)
waterData$FecalColiform = as.numeric(waterData$FecalColiform)
waterData$TotalColiform = as.numeric(waterData$TotalColiform)
sum(is.na(waterData$DO))
sum(is.na(waterData$Temp))
sum(is.na(waterData$pH))
sum(is.na(waterData$BOD))
sum(is.na(waterData$Conductivity))
sum(is.na(waterData$FecalColiform))
sum(is.na(waterData$TotalColiform))
waterData$DO = replace(waterData$DO, is.na(waterData$DO), mean(waterData$DO))
waterData$pH = replace(waterData$pH, is.na(waterData$pH), mean(waterData$pH))
waterData$Temp = replace(waterData$Temp, is.na(waterData$Temp), mean(waterData$Temp))
waterData$BOD = replace(waterData$BOD, is.na(waterData$BOD), mean(waterData$BOD))
waterData$Conductivity = replace(waterData$Conductivity, is.na(waterData$Conductivity), mean(waterData$Conductivity))
waterData$FecalColiform = replace(waterData$FecalColiform, is.na(waterData$FecalColiform), mean(waterData$FecalColiform))
waterData$TotalColiform = replace(waterData$TotalColiform, is.na(waterData$TotalColiform), mean(waterData$TotalColiform))
max(waterData$pH)
waterData$pH = replace(waterData$pH, waterData$pH>88, 88.13)
```



```{r}
cropData = read.csv("crop_production.csv")
str(cropData)
```

```{r}
summary(lm(Production~Area, cropData))
```

```{r}
waterStates = unique(waterData$STATE)
waterStates
```



```{r}
cropData$State_Name = sub("Dadra and Nagar Haveli","Daman and Diu",cropData$State_Name)
cropStates = unique(cropData$State_Name)
print(cropStates)
```

```{r}
waterYears = unique(waterData$year)
cropYears = unique(cropData$Crop_Year)
print(waterYears)
print(cropYears)
cropData = cropData %>%
  filter(Crop_Year %in% 2003:2014)
cropYears = unique(cropData$Crop_Year)
print(cropYears)

```

```{r}
sum(is.na(waterData))
nrow(waterData)
water_data = na.omit(waterData)
nrow(water_data)
sum(is.nan(water_data$FECAL.COLIFORM..MPN.100ml.))
```

```{r}
#unique(water_data$STATE)
k=c('Telangana','Andhra Pradesh',"AP",'A.P.','Arunachal Pradesh','Assam','Bihar','Chhattisgarh','Goa', 'Gujarat','Haryana', 'Himachal Pradesh','H.P.','Jammu and Kashmir', 'Jharkhand','Karnataka','Kerala','Madhya Pradesh','Maharashtra', 'Manipur','Meghalaya','Mizoram','Nagaland','Odisha','ORISSA','Punjab', 'Rajasthan', 'Sikkim','Tamil Nadu', 'Tamilnadu','Tripura','Uttar Pradesh','Uttarakhand','West Bengal','Andaman and Nicobar',
'Chandigarh','Dadra and Nagar Haveli','Daman and Diu','Daman','Lakshadweep','Puducherry','PONDICHERRY','Delhi')
length(k)
k <- toupper(k)
pattern = paste(paste(".*\\b(", paste(k, collapse="|")), ")\\b.*")
pattern
```

```{r}
water_data$states =  sub(pattern, "\\1", water_data$STATE)
#water_data$states =  sub(pattern, "\\1", water_data$STATE)
water_data$states =  ifelse(water_data$states == "NAN", sub(pattern, "\\1", water_data$LOCATIONS), water_data$states)
water_data$states = sub('^(A.P.|AP)$','ANDHRA PRADESH',water_data$states)
water_data$states = sub("^H.P.$","HIMACHAL PRADESH",water_data$states)
water_data$states = sub("^TAMILNADU$","TAMIL NADU",water_data$states)
water_data$states = sub("^DAMAN$","DAMAN AND DIU",water_data$states)
water_data$states = sub("^PONDICHERRY$","PUDUCHERRY",water_data$states)
water_data$states = sub("^tripura$","TRIPURA",water_data$states)
water_data$states = sub("^ORISSA$","ODISHA",water_data$states)
#View(water_data)
unique(water_data$states)
```


```{r}
#install.packages("stringr")
library(stringr)
print(paste("Number of unique sates generated : ",length(unique(water_data[str_detect(water_data$states, pattern),]$states))))
unique(water_data[str_detect(water_data$states, pattern),]$states)
# number of states matching with dataset
nrow(water_data[str_detect(water_data$states, pattern),])
# total rows (1679-1406 = 273 yet to be modified)
nrow(water_data)
```

```{r}
# arranging the data frame
#View(water_data)
water_data_clean = water_data[str_detect(water_data$states, pattern),][,-c(1,2,3)]
water_data_clean = water_data_clean[,c(9,10,1,2,3,4,5,6,7,8)]
water_data_clean <- rename(water_data_clean,STATE=states)
#View(water_data_clean)
```

```{r}
nrow(cropData)
sum(is.na(cropData))
cropData = na.omit(cropData)
```


#Extracted data 
```{r}
sWater_data <- water_data_clean
nrow(sWater_data)
#View(sWater_data)
years = unique(waterData$year)
states <- tolower(unique(water_data[str_detect(water_data$states, pattern),]$states))
req_Crop_data <- cropData %>%
  filter(tolower(State_Name) %in% states)
nrow(req_Crop_data)
summary(lm(Production~Area, cropData))
```

```{r}
sCrop_data = req_Crop_data %>% group_by(Crop_Year, State_Name) %>%
  summarize(totalArea=mean(Area), totalProd=mean(Production),.groups = 'drop')
sWater_data = sWater_data %>% group_by(year, STATE) %>%
  summarize(Temp=mean(Temp), DO=mean(DO), pH = mean(pH),Conductivity=mean(Conductivity), BOD = mean(BOD), Nitrates=mean(Nitrates), FecalColiform = mean(FecalColiform), TotalColiform = mean(TotalColiform), .groups = 'drop' )

```


# Merging data
```{r}
sCrop_data <- rename(sCrop_data,STATE=State_Name)
sCrop_data$STATE <- tolower(sCrop_data$STATE)
sWater_data$STATE <- tolower(sWater_data$STATE)

sCrop_data = rename(sCrop_data, year=Crop_Year)
View(sCrop_data)
View(sWater_data)
com_data <- merge(sWater_data,sCrop_data,by=c("year","STATE"))
View(com_data)
```

```{r}
nrow(com_data)
#View(waterData)
```

# Basic relation ship analisys
```{r}
AreaRel = lm(totalProd~totalArea, com_data)
summary(AreaRel)
```
```{r}
crops = unique(req_Crop_data$Crop)
print(crops)
which.max(table(req_Crop_data$Crop))
```

```{r}
cropWiseRel = function(crop){
  data = req_Crop_data %>%
    filter(Crop == crop)
  data = data %>% group_by(Crop_Year, State_Name) %>%
    summarize(totalArea=mean(Area), totalProd=mean(Production),  .groups = 'drop')
  data <- rename(data,STATE=State_Name, year=Crop_Year)
  data$STATE <- tolower(data$STATE)
  #View(data)
  com_data <- merge(sWater_data,data,by=c("year","STATE"))
  if (nrow(com_data)!=0){
    DS_AreaRel = lm(totalProd~totalArea, com_data)
    #summary(DS_AreaRel)
    if(is.na(DS_AreaRel$coefficients[2])==F && DS_AreaRel$coefficients[1] != 0){
      cat("\n",crop,":\n")
      print(DS_AreaRel$coefficients)
      cat("correlation : ", cor(com_data$totalArea, com_data$totalProd),"\n")
    }
  }
}
```

# crop wise relationship of the total production by total area provided
```{r}
for(i in crops){
  cropWiseRel(i)
}
```

# State wise yield 
```{r}
states = unique(com_data$STATE)

# state-wise relation of the yield with respect to the each containment
statewise <- function(bigdata,yieldtype,type){
  for(i in states){
    data=bigdata %>% filter(STATE==i)
    if(nrow(data)!=0){
      # capping yield attribute for each state
      qnt <- quantile(data$yield, probs=c(.25, .75), na.rm = T)
      caps <- quantile(data$yield, probs=c(0, 1), na.rm = T)
      H <- 1.5 * IQR(data$yield)
      data$yield[data$yield < (qnt[1] - H)] <- caps[1]
      data$yield[data$yield > (qnt[2] + H)] <- caps[2]

      # capping water chemical attribute for each state
      qnt <- quantile(data$yieldtype, probs=c(.25, .75), na.rm = T)
      caps <- quantile(data$yieldtype, probs=c(0, 1), na.rm = T)
      H <- 1.5 * IQR(data$yieldtype)
      data[yieldtype][data[yieldtype] < (qnt[1] - H)] <- caps[1]
      data[yieldtype][data[yieldtype] > (qnt[2] + H)] <- caps[2]
      
      Rel = lm(paste0(yieldtype,'~',type),data)
      if(is.na(Rel$coefficients[2]) == F || is.na(Rel$coefficients[1] == F)){
        cat(i,":\n")
        print(Rel$coefficients)
        cat("correlation : ", cor(data$yield,data[yieldtype]),"\n\n")
      }
    }
  }
}

# total increase in production by summing for each state with respect to each containment
statewises = function(bigdata,yieldtype,para){
  for(i in states){
    data = bigdata %>% filter(STATE==i)
    data = summarize(data,totalyield=sum(yield), typeyield=sum({{yieldtype}}))
    # or
    # data = summarize(data,totalyield=sum(yield), typeyield=sum(data$yieldtype))
    cat(i,":\n")
    cat("Total_yield            : ",data$totalyield,"\n")
    cat(para,"_total_yield  : ",data$typeyield,"\n\n")
  }
}
```

#Relation ship between production and water qualty parameters
```{r}
com_data$yield = com_data$totalProd/com_data$totalArea
```

## Nitrate
```{r}
com_data$Nitrates_yield = ifelse(com_data$Nitrates==0,com_data$totalProd/com_data$totalArea,
                                 com_data$totalProd/(com_data$totalArea*com_data$Nitrates))
statewise(com_data,'Nitrates_yield','Nitrates')
statewises(com_data,Nitrates_yield,"Nitrates")
```

## DO
```{r}
com_data$DO_yield = ifelse(com_data$DO==0,com_data$totalProd/com_data$totalArea,(com_data$totalProd*com_data$DO)/com_data$totalArea)
statewise(com_data,'DO_yield','DO')
statewises(com_data,DO_yield,"DO")
```

## BOD
```{r}
com_data$BOD_yield = ifelse(com_data$BOD==0,com_data$totalProd/com_data$totalArea,com_data$totalProd/(com_data$totalArea*com_data$BOD))
statewise(com_data,'BOD_yield','BOD')
statewises(com_data,BOD_yield,"BOD")
```

## FecalColiform
```{r}
com_data$FecalColiform_yield = ifelse(com_data$FecalColiform==0,com_data$totalProd/com_data$totalArea,com_data$totalProd/(com_data$totalArea*com_data$FecalColiform))
statewise(com_data,'FecalColiform_yield','FecalColiform')
statewises(com_data,FecalColiform_yield,"FecalColiform")
```

## TotalColiform
```{r}
com_data$TotalColiform_yield = ifelse(com_data$TotalColiform==0,com_data$totalProd/com_data$totalArea,com_data$totalProd/(com_data$totalArea*com_data$TotalColiform))
statewise(com_data,'TotalColiform_yield','TotalColiform')
statewises(com_data,TotalColiform,"TotalColiform")
```
# Random Forest
```{r}
# install.packages("caret", dependencies = TRUE)
# install.packages("randomForest")
library(caret)
library(randomForest)
library(MLmetrics)
```

```{r}
set.seed(51)
MAE=c()
MSE=c()
RMSE=c()
R2=c()
states = unique(com_data$STATE)
statewiseRF = function(){
  for(i in states){
    state = com_data %>%
      filter(STATE == i)
    if(nrow(state)>1){
      model <- train(yield ~ Temp + DO + pH + Conductivity + BOD + Nitrates + FecalColiform + TotalColiform, state,'rf' 
                     ,trControl = trainControl(method = 'cv',number = 4))
      pred_yeild = predict(model, com_data[-c(1,2,11,12,13)])
      fmae = MAE(pred_yeild, com_data$yield)
      fmse = MSE(pred_yeild, com_data$yield)
      frmse = RMSE(pred_yeild, com_data$yield)
      fr2 = R2(pred_yeild, com_data$yield)
      MAE <<- append(MAE,fmae)
      MSE <<- append(MSE,fmse)
      RMSE <<- append(RMSE,frmse)
      R2 <<- append(R2,fr2)
      cat(i,":\n")
      cat(" MAE:", fmae, "\n", "MSE:", fmse, "\n", "RMSE:", frmse, "\n", "R-squared:", fr2, "\n")
    }
  }
}
# qnt <- quantile(com_data$yield, probs=c(.25, .75), na.rm = T)
# caps <- quantile(com_data$yield, probs=c(0, 0.75), na.rm = T)
# H <- 1.5 * IQR(com_data$yield)
# com_data$yield[com_data$yield < (qnt[1] - H)] <- caps[1]
# com_data$yield[com_data$yield > (qnt[2] + H)] <- caps[2]
# boxplot(com_data$yield,com_data$TotalColiform)
# summary(com_data$yield)
options(warn=-1)
statewiseRF()
options(warn=0)
```

```{r fig.width = 13}
ggplot(data.frame(states[1:14],MAE), aes(states[1:14],MAE)) + geom_point() + geom_line(aes(group = 1))
ggplot(data.frame(states[1:14],MSE), aes(states[1:14],MSE)) + geom_point() + geom_line(aes(group = 1))
ggplot(data.frame(states[1:14],RMSE), aes(states[1:14],RMSE)) + geom_point() + geom_line(aes(group = 1))
ggplot(data.frame(states[1:14],R2), aes(states[1:14],R2)) + geom_point() + geom_line(aes(group = 1))
```


## Usage of neural Network over Combined data

```{r}
# install.packages("neuralnet")
library(neuralnet)
```

## for scaling the data and accuracy calculation
```{r}
library(caret)
#install.packages("MLmetrics")
library(MLmetrics)
# install.packages("Metrics")
```

## Functions for neural network analysis
```{r}
scale = function(s){
  mod = preProcess(data.frame(s), method = 'range')
  scale = predict(mod, data.frame(s))
  scale = scale$s
}
meanSqErr = function(a,b){
  suM = 0
  for(i in seq(length(a))){
    t = (a[i]-b[i,])^2
    suM = suM + t
  }
  return (suM/length(a))
}
stateWiseNNAcc = function(s){
  stateData = com_data %>%
  filter(STATE == s)
  stateData$totalArea = scale(stateData$totalArea)
  stateData$Temp = scale(stateData$Temp)
  stateData$DO = scale(stateData$DO)
  stateData$pH = scale(stateData$pH)
  stateData$Conductivity = scale(stateData$Conductivity)
  stateData$BOD = scale(stateData$BOD)
  stateData$Nitrates = scale(stateData$Nitrates)
  stateData$FecalColiform = scale(stateData$FecalColiform)
  stateData$TotalColiform = scale(stateData$TotalColiform)
  stateData$totalProd = scale(stateData$totalProd)
  # op_Sca = preProcess(data.frame(stateData$totalProd), method = "range")
  # stateData$totalProd = as.numeric(predict(op_Sca, data.frame(stateData$totalProd))$stateData.totalProd)
  # View(stateData)
  par_ProRel = neuralnet(totalProd~totalArea+Temp+DO+pH+Conductivity+BOD+Nitrates+FecalColiform+TotalColiform, stateData, stepmax = 1e+08)
  
  res = predict(par_ProRel, stateData[c(-12)])
  print(paste(paste(paste("RMSE in state ", s), " is "), meanSqErr(res, stateData[c(12)])))
  return(par_ProRel)
}
```

```{r}
# mses = c()
unique(com_data$STATE)
# summary(stateWiseNNAcc("tamil nadu"))
# plot(stateWiseNNAcc("tamil nadu"))
```


# Analysis using the weights generated by neural network, for each state

```{r}
plot(stateWiseNNAcc("gujarat"))
```

## Inference : In gujarat it can be seen that, increase in the temperature, Nitrates is leading to decrease in total production, while increase in the DO, BOD is leading to the increase in the total crop production  


```{r}
plot(stateWiseNNAcc("karnataka"))
```
## Inference : In karnataka it can be seen that, increase in the DO,pH,Conductivity,FecalColiform is leading to decrease in total production, while increase in the BOD,NItrates is leading to the increase in the total crop production  

```{r}
plot(stateWiseNNAcc("kerala"))
```

## Inference : In kerala it can be seen that, increase in the DO,pH,BOD, Notrates and TotalColiform is leading to decrease in total production, while increase in the temperature, FecalColiform is leading to the increase in the total crop production 

```{r}
plot(stateWiseNNAcc("maharashtra"))
```
## Inference : In maharastra it can be seen that, increase in the Temperature,DO,pH,Noterates and TotalColiform is leading to decrease in total production, while increase in the BOD, FecalColiform is leading to the increase in the total crop production 

```{r}
plot(stateWiseNNAcc("odisha"))
```

## Inference : In odisha it can be seen that, increase in the temperatue,DO,pH,BOD,FecalColiform is leading to decrease in total production, while increase in the TotalColiform is leading to the increase in the total crop production 

```{r}
plot(stateWiseNNAcc("rajasthan"))
```

## Inference : In rajasthan it can be seen that, increase in most of the variables is leading to the increase in the total crop production 

```{r}
plot(stateWiseNNAcc("madhya pradesh"))
```

## Inference : In Madhya Pradesh it can be seen that, increase in the Temperature,DO,pH,Nitrates is leading to decrease in total production, while increase in the Conductivity,BOD, TotalColiform is leading to the increase in the total crop production 

```{r}
plot(stateWiseNNAcc("tamil nadu"))
```

## Inference : In Tamil Nadu it can be seen that, increase in the DO,pH,BOD,TotalColiform is leading to decrease in total production, while increase in the Temperature,Conductivity,Niterates, is leading to the increase in the total crop production

```{r}
plot(stateWiseNNAcc("andhra pradesh"))
```

## Inference : In ANdhra Pradesh it can be seen that increase in the Temperature,BOD,TotalColiform is leading to decrease in total production, while increase in the DO,pH,Conductivity,Niterates,FecalColiform is leading to the increase in the total crop production

```{r}
plot(stateWiseNNAcc("goa"))
```

## Inference : In Goa it can be seen that increase in the FecalColiform,BOD is leading to decrease in total production, while increase in the DO,TotalColiform is leading to the increase in the total crop production

```{r}
plot(stateWiseNNAcc("himachal pradesh"))
```

## Inference : In Himachal Pradesh it can be seen that increase in the pH,BOD,Niterates is leading to decrease in total production, while increase in the FecalColiform,TotalColiform is leading to the increase in the total crop production

```{r}
plot(stateWiseNNAcc("meghalaya"))
```

## Inference : In Meghalaya it can be seen that, increase in the Temperature,DO,BOD,Niterates,FecalColiform,TotalCoiform is leading to the increase in the total crop production

```{r}
plot(stateWiseNNAcc("punjab"))
```

## Inference : In Punjab it can be seen that increase in the DO,Conductivity,TotalColiform is leading to decrease in total production, while increase in the Temperature,Nitrates,FecalColiform is leading to the increase in the total crop production

```{r}
plot(stateWiseNNAcc("tripura"))
```

## Inference : In Tripura it can be seen that increase in the Nitrates,FecalColiform,DO is leading to decrease in total production, while increase in the pH,Conductivity,BOD is leading to the increase in the total crop production

```{r}
plot(stateWiseNNAcc("daman and diu"))
```

## Inference : In Daman and Diu it can be seen that increase in the DO,Nitrates,FecalColiform,TotalColiform is leading to decrease in total production, while increase in the BOD is leading to the increase in the total crop production

```{r}
plot(stateWiseNNAcc("haryana"))
```

## Inference : In Haryana it can be seen that, increase in the DO,pH,BOD is leading to decrease in total production

# It can be be seen that in some of the states, the totalArea had a negative impact on the total production, which is due to the diversity of the crops information present in each state, this can be resolved by analysing the information of each crop in each state, which is the future scope for this project.We had implemented Random forest algorithm as well for better understanding of the behavior of chemical on the crop, which is given below.

